A Large and Variable Leading Tail of Helium in a Hot Saturn Undergoing Runaway Inflation

### 2023-12-04 paper/paper1/ms.tex

Figure \ref{fig:KeplerianShear} shows an illustration of Keplerian shear adapted to the system properties of \hatpb.  In this shear-dominated scenario, the planetary wind launches primarily from the dayside, with relatively little or no wind launched from the nightside.  The planet wind initially launches radially outward from the exobase, with the strongest wind located near the sub-stellar point, the line connecting the planet to the star along the vertical axis in the figure.  Inefficient grazing incidence heating near the terminator subdues the mass loss in the $x-$direction, meaning that this wind exhibits not a hemispherical shape, but more concentration along the star-planet line.  \edit1{The gas increasingly experiences Keplerian shear past the Roche lobe, manifesting as prograde acceleration.  In the rotating frame, the Coriolis effect causes the inward-moving gas to advance ahead of the planet. In the inertial frame, the gas is launched onto an orbit with a smaller periapse than the planet, and so runs ahead.  The accelerating flow} eventually overtakes the planet completely, with the prospect of extending to hundreds of planetary radii.  The mass loss has to be large enough to shield the material to make it observable in the metastable \ion{He}{1} 10833 \AA~triplet.

\begin{figure}
  \includegraphics[width=0.8\linewidth]{KeplerianShear_v0p5_alt.png}
  \caption{Schematic of Keplerian orbital shear.  Locations outside the Roche lobe experience orbital shear from the Keplerian potential.  A wind launched primarily from the dayside will tend to form a leading tail.}
  \label{fig:KeplerianShear}
\end{figure}

>>>>

Figure \ref{fig:KeplerianShear} shows an illustration of Keplerian shear adapted to the system properties of \hatpb.  In this shear-dominated scenario, the planetary wind launches primarily from the dayside, with relatively little or no wind launched from the nightside.  The planet wind initially launches radially outward from the exobase, with the strongest wind located near the sub-stellar point, the line connecting the planet to the star along the vertical axis in the figure.  Inefficient grazing incidence heating near the terminator subdues the mass loss in the $x-$direction, meaning that this wind exhibits concentration along the star-planet line.  \edit1{The gas increasingly experiences Keplerian shear past the Roche lobe, manifesting as prograde acceleration.  In the rotating frame, the Coriolis effect causes the inward-moving gas to advance ahead of the planet. In the inertial frame, the gas is launched onto an orbit with a smaller periapse than the planet, and so runs ahead.  The accelerating flow} rapidly overtakes the planet completely, with the prospect of extending to hundreds of planetary radii.
\edit1{The overview of the orbit in Figure \ref{fig:KeplerianShear} shows one conceivable path of the gas motion.  The exact extent of the path depends on several unknowns.  For example, } the column density has to be large enough to shield the material to make it observable in the metastable \ion{He}{1} 10833 \AA~triplet.

\begin{figure*}
  \centering
  \includegraphics[width=0.95\textwidth]{Fig14_KeplerianShear_revised.png}
  \caption{\edit1{Top-down illustration of mass loss from the HAT-P-67~b system.  A planetary wind launched primarily from the dayside will tend to form a leading tail.  The strongest dayside mass loss will occur at the sub-solar point, with a fall off towards zero near the planet terminator.  The small Roche lobe of merely $<2.7R_p$ means that gas promptly experiences Keplerian orbital shear after launching past the L1 point.  The orbital overview depicts a path of the mass loss initially towards the star, decelerating into the headwind of the stellar wind, and then reversing direction outward in a blueshift. The observed redshift and blueshift of any given sightline will depend on the Helium ionization fraction, which depends on the shielding of gas along the line of sight.} }
  \label{fig:KeplerianShear}
\end{figure*}
